---
title: "Assignment 2"
author: "Marguerite CAMERON"
date: "13 octobre 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r}
source("functions.R")
```

```{r}
sum_column <- sum(iris$Sepal.Width)
  x <- iris[["Sepal.Width"]] 
  if (is.null(x))
    return <- NULL
```

```{r}
sum_column <- sum(iris$Sepal.Width)
print(sum_column)
```

# Other sum function
I tried to find other ways to make a sum function that did not use the default, but was not at all successful.  I tried to use a cumulative function.  I tried to make a list of all the members of a vector, thinking I could add them together after, and I could not make that work either.  

I did not get around to trying the division function.  


```{r}
library(ggplot2)
theme_set(theme_bw())

# plot
p <- ggplot(iris, aes(Species, Sepal.Width))
p + geom_violin() + 
  labs(title="Violin plot", 
       subtitle="Species vs Sepal Wdith",
       caption="Source: iris",
       x="species of iris",
       y="sepal width in cm")
```

```{r}
d = iris
var = "Sepal.Length"
grouping_var = "Species"
grouped_violin_plot <- function(d, var, grouping_var) 
# Create the base ggplot object
p <- ggplot2::ggplot(d, ggplot2::aes_string(y=var,
x=grouping_var,
fill=grouping_var))
p1 <- p + geom_violin()
return(p1)

print(grouped_violin_plot(iris, "Sepal.Length", "Species"))
```


```{r}
d = iris
var = "Sepal.Length"
grouping_var = "Species"
grouped_violin_plot <- function(d, var, grouping_var) {
# Create the base ggplot object
p <- ggplot2::ggplot(d, ggplot2::aes_string(y=var,
x=grouping_var,
fill=grouping_var))
p1 <- p + geom_violin() + labs(title="Iris Data")
return(p1)
}
print(grouped_violin_plot(iris, "Sepal.Length", "Species"))

```

# Changing the color
I tried to use the "scale_color_manual" function like it shows on the online tutorial, naming the species categories and assigning them colors, but it just produced exactly the same graph.  I also tried to use a couple other examples I found online, and did not succeed in having an effect. 

```{r}
 d = iris
  var = "Sepal.Length"
  grouping_var = "Species"
  group1 = "versicolor"
  group2 = "viriginica"
  difference_in_medians <- function(d, var, grouping_var, group1, group2) 
    d_1 <- dplyr::filter(d, get(grouping_var) == group1)
    d_2 <- dplyr::filter(d, get(grouping_var) == group2)
    # YOUR CODE HERE: assign the difference in the medians to to the variable 'result'
    return(result)
```
# difference in medians
My console says that bindrcpp was built under R version 3.4.2 and would not run the function.  I tried to install the bindrcpp package, and instead of putting it into R like other packages, the program put it somewhere in a temp folder, and I was not able to find it (using search in Windows explorer).


```{r}
randomize_rts <- function(iris) {
result <- d
n <- nrow(d)
d$logRT <- sample(d$logRT, n)
return(d)
}
```

```{r}
# Randomize the order of a column.
#
# ARGUMENTS:
# d: a data frame or tibble
# var: the name of a column of d containing the variable to randomize,
# provided as a string
#
# RETURN VALUE:
# A data frame or tibble exactly the same as d, except with the order of
# var permuted randomly.
#
d = iris
var = "Sepal.Length"

randomize <- function(d, var) {
d[[var]] <- sample(var)
return(d)
}
```

```{r}
iris$Sepal.Width[1:10]
if(!exists(".Random.seed")) set.seed(NULL)
previous_seed <- .Random.seed
set.seed(1)
randomize(iris, "Sepal.Width")$Sepal.Width[1:10]
randomize(iris, "Species")$Species[1:10]
randomize(iris, "Species")$Sepal.Width[1:10]
set.seed(previous_seed)
```

```{r}

# Perform a permutation test for two groups.
#
# ARGUMENTS:
# d: a data frame or tibble
# var: the name of the column in d on which the test statistic will be calculated,
# provided as a string
# grouping_var: the name of the column in d which gives the grouping
# group1: the value of grouping_var corresponding to the first group
# group2: the value of grouping_var corresponding to the second group
# statistic: a function yielding a test statistic, which takes as input
# a data frame, the name of a variable on which to calculate the
# test statistic, the name of a grouping variable, the value of
# the grouping variable corresponding to the first group, and
# the value of the grouping variable corresponding to the second
# group
# n_samples: the number of permutation samples to draw (default: 9999)
#
# RETURN VALUE:
#
# A list containing two elements:
#
# - observed: the value of statistic() in d
# - permuted: a vector containing the values of statistic() under n_samples
# permutations
#
d = iris
var = "Sepal.Length"
grouping_var = "Species"
group1 = "versicolor"
group2 = "virginica"
permutation_twogroups <- function(d, var, grouping_var, group1, group2, statistic,
n_samples=9999) {
observed_statistic <- statistic(d, var, grouping_var, group1, group2)
permutation_statistics <- rep(0, n_samples)
for (i in 1:n_samples) {
# YOUR CODE HERE: use randomize(...) to create a permutation and then
# fill in the vector permutation_statistics with the
# value of statistic(...) for this new permutation
}
result <- list(observed=observed_statistic,
permuted=permutation_statistics)
return(result)
}

if(!exists(".Random.seed")) set.seed(NULL)
previous_seed <- .Random.seed
set.seed(1)
ptest_1 <- permutation_twogroups(iris, "Sepal.Width", "Species", "versicolor",
"virginica", difference_in_medians, n_samples=10)
ptest_2 <- permutation_twogroups(iris, "Sepal.Width", "Species", "versicolor",
"virginica", difference_in_medians, n_samples=10)
ptest_3 <- permutation_twogroups(randomize(iris, "Sepal.Width"), "Sepal.Width",
"Species", "versicolor", "virginica",
difference_in_medians, n_samples=10)
set.seed(previous_seed)
print(ptest_1)
print(ptest_2)
print(ptest_3)
print(ptest_3$observed)
print(ptest_3[["observed"]])
```

# Exercise 3d
Does it matter whether you permute var or grouping_var?

Yes.  var is a vector containing a range of different values.  Therefore, the permutation result can change from one permutation to another.  grouping_var only contains a few available categories in this case : the handful of iris species in the data base.  If we permuted the grouping variable, there would be less variation in the result, since it only has a few values to permute.  The sampling variety is much greater getting a random sample from a vector that has a better variety of values.  


```{r, cache=T}
if(!exists(".Random.seed")) set.seed(NULL)
previous_seed <- .Random.seed
set.seed(1)
ptest <- permutation_twogroups(iris, "Sepal.Width", "Species", "versicolor",
"virginica", difference_in_medians)
set.seed(previous_seed)
```



